\LARGE \bf \pyRobots{}, a toolset for robot executive control

### 2014-09-18 paper.tex
Lot of little things

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\pyRobots}{\textsc{pyRobots}\ }

\title{\LARGE \bf
    \pyRobots, a toolkit for robot executive control
}

@@

Presented is \pyRobots, a new software toolkit for the executive control of
complex robotic systems. Borrowing ideas from previous tools like {\sc
urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
develop (pre-emptive) concurrent and event-based executive controllers.

Designed in a \emph{bottom-up} fashion, out of the need for a practical,
unobstrusive tool suitable for rapid prototyping of complex interaction
scenarios, \pyRobots also exposes several simple yet convenient abstractions for
physical resources management and pose representation.

@@

Orchestrating the activity of a robot is a central issue of robotics: \emph{what
to do when?} Traditionally, in the famous 3-layer architectures~\fixme{ref?},
this function is implemented in the so-called \emph{deliberative layer} where
decisions are made based on perceptions and on the current internal state, and
executed by sending orders to a \emph{functional layer}.

@@

This article introduces our attempt at designing an unobtrusive execution control
toolkit, that aims at addressing the \emph{practicality} issue: it has been
designed and implemented from the bottom-up, starting from actual needs when
running complex robots in largely unpredictable scenarios (typically,
loosely constrained human-robot interactions), and in the real world.

Instead of an \emph{environment} or a \emph{framework}, which would suggest
strong design and development constraints, \pyRobots can be conceived as a set of
powerful software helpers to write parallel, event-based, high-level robot
controllers.

It is written for and in Python, which ensure familiarity, fast development
cycles and broad support for interfacing with existing robots and middlewares.
This must be emphasized: \pyRobots is \emph{not} a middleware. It purposefully
provides no mechanisms to connect to other modules or components. It instead
\emph{uses} one (or several) middleware to actually communicate with the robot
and perform actions.

One the other hand, \pyRobots does not provide formal models or verifiability:
in its current state, it does not attempt to contribute in this field. It
focuses instead on providing unobtrusive tools to control robots in environments
that require carrying out many tasks in parallel, in an unpredictable,
event-based manner.

@@

Any robotic experiment requires a certain amount of high-level control to
supervise the robot behaviour, and, expectedly, many projects have attempted to
design and provide tools for this task.  We focus this review on \emph{tools
dedicated to the explicit implementation of high-level behaviours}, excluding
complete control stacks (that would typically include a middleware and hardware
abstractions), meta-tools (like component generators or model checkers) and
(semi-)automatic techniques (like learning by demonstration).  In contrast with
middlewares for which the robotic community relies on a few \emph{de-facto}
standards (ROS, YARP, OpenRTM), we want to underline that no such widely-used
softwares exist for execution control, and we believe that most of today's robot
behaviours are implemented as \emph{ad-hoc} scripts.

@@

Domain-specific languages (DSL), in the widest sense, represent the largest body
of literature on high-level behaviour programming. They can either be ``actual''
domain-specific languages like PRS~\cite{ingrand1996prs},
RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
derivative like {\sc ReadyLOG}~\cite{ferrein2007robot} or, more recently,
URBI~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
languages like Python or Ruby, or logic languages like Prolog) to provide
high-level paradigms and control structures well suited to the programming of
robots. {\sc cram}~\cite{beetz2010cram} (based on Lisp and Prolog), {\sc
roby}~\cite{joyeux2009plan} (based on Ruby), {\sc
teer}~\footnote{\url{http://wiki.ros.org/executive_teer}} (based on Python), are
examples of such eDSL.

@@

\pyRobots positions itself in this landscape as a lightweight eDSL (in a similar
fashion to {\sc roby} or {\sc teer}), that focuses on high-level behaviours
implementation. It seeks unobstrusive interaction with the other components of
the robot software stack, using Python as ubiquitous binding language. Similar
to URBI, it focuses on transparent preemptive concurrency and event-based programming.
Also, \pyRobots' model favours hierarchical task composition over plans, and does not
provide by itself any task planning service, delegating this to external tools.

@@

Before introducing and explaining step-by-step a complete working example in the
next section, it is useful to first introduce the main concepts of \pyRobots,
and provide an overview of the features and mechanisms. Since \pyRobots is an
open-source software, the interested reader can also refer to the source code,
available here:~\url{https://github.com/chili-epfl/pyrobots}.

\begin{itemize}
    \item Lock specific resources with a simple \python|@lock(...)| in front of the
        actions. When starting, actions will wait for resources to be
        available if needed.
    \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
    \item Poses are managed explicitly and can easily be
        transformed from one reference frame to another one
        (integrates with ROS TF when available).
    \item Extensive logging support to debug and replay
        experiments.
\end{itemize}

@@

\pyRobots core concepts are simple: on one hand, the user creates a
\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
different low-level \textbf{controllers} (proxies to actuators, typically
provided by a middleware like ROS) and the \textbf{state} of the robot
(typically a set proxies to the sensors, or a connector to a database, etc.,
accessed through a dictionary-like object).

On the other hand, the user creates as many high-level \textbf{actions} as
desired, as regular Python functions. Actions are automatically added to the
\textbf{robot}, and can access its state and controllers to perform actual
physical actions.

@@

Finally, the user can declare arbitrary \textbf{resources}, usually
corresponding to the robot components (for instance, the wheels, the camera,...)
and guarantee exclusive access to the required resource to specific actions.

@@

\emph{Actions} in \pyRobots are always asynchronous. They are implemented as
\emph{futures} (also known as \emph{promises}) and therefore executed in
independent threads. To the developer, however, they appear as regular Python
function, simply annotated with the \python|@action| decorator, and are invoked
like any other function.  Since they are asynchronous, these functions may
perfectly block or even never return (typically useful to implement lasting
background behaviours).

@@

\paragraph{Signaling} Behaviours often need to be paused of cancelled.  This
raises an issue that needs to be addressed in robotics: before suspending an
action, one often wants to first bring back the robot into some form of rest
state (\eg one would not want to suddenly cancel a \python|walk| action in the
middle of a step without first bringing back the foot to the ground).

\pyRobots addresses this issue through signals, and extends standard
Python thread into \emph{signaling threads}, relying on Python's usual syntax
for exception handling. Listing~\ref{lst:signals}
illustrates this mechanism. The action is instantiated and started at line 8,
and cancelled at line 10. This send a \python|ActionCancelled| signal to the
action that the user can handle using an \python|except| statement.

\begin{listing}[H]
\begin{pythoncode}
    @action
    def my_action(robot):
      try:
        robot.do_something_long()
      except ActionCancelled:
        robot.go_back_to_rest_pose()

    action = robot.my_action()
    time.sleep(1)
    action.cancel()
\end{pythoncode}
\caption{Handling a cancellation signal}
\label{lst:signals}
\end{listing}

Pausing actions is more complex (the action internal state must be saved, a mean
to resume the action where it was interrupted must be provided, the user must be
given ways to code transition behaviours when pausing and resuming) and is not
currently supported by \pyRobots.

@@

Sharing the same syntax with URBI, events in \pyRobots can by monitored through
the methods \python|on| (one shot event) and \python|whenever| (continuous event
monitoring). Event conditions can be either a predicate (that must take as
single argument the robot instance) or a pair of $\{\text{state member},
\text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$. An
example of event usage can be seen at line 67 of listing~\ref{mwe}.

When the event condition is verified, the callback (usually, an action) is
invoked from the event monitor thread.

@@

Usually, physical robot resources (mainly actuators, but possibly sensors as
well) require exclusive use (it is seldom desirable that two different actions
control the wheels of a robot at the same time).

To leave full control regarding the required resource modeling granularity, the
programmer define them himself, either as stand-alone resources or as compound
resources (listing~\ref{lst:resources}).

@@

    robot.lift_box()
    robot.wave_hand() # waits until lift_box is over
    robot.scratch_head() # skipped if lift_box 
                         # is still running

\end{pythoncode}
\caption{Resource usage is defined at the action-level, through annotations.}
\label{lst:resources}
\end{listing}

The user then annotates the actions with the (possibly multiple) resources they
require (lines 6, 11 and 16 of listing~\ref{lst:resources}). When executed, the
action acquires a lock on the resource (and all dependent resources in the case
of a compound resource) which is released upon action completion. In the
meantime, other actions requiring the same resource will either wait until the
resource is release (line 21) or be skipped (line 22) if the \python|@lock|
annotation specifies not to wait (line 16).

@@

\pyRobots offers a similar approach \emph{within} the control script, adding a
pinch of syntactic sugar to make poses and frame management transparent for the
high-level action controller.

As illustrated in listing~\ref{lst:poses}, poses can be flexibly defined by a
simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt map} is then
assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y, z, rx, ry,
rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx, ry, rz,
frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x, y, z,
qx, qy, qz, qw, frame\}$ or by a dictionary that defines any of these 8 values.
Internally, \pyRobots normalizes and stores the poses in their full 8-tuple
expansion.

@@

\end{pythoncode*}
\caption{Examples of pose normalization.}
\label{lst:poses}
\end{listing}

@@

These transformations are directly computed, but \pyRobots also takes advantage
of middleware-specific infrastructure (when available) to provide known frames.
ROS TF, Aldebaran's {\tt NAOqi} frame conventions, etc. are used when available
through specializations of the \pyRobots \python|PoseManager| class.

@@

Running robots in real-world, \ie dynamic and unpredictable, environments
entails repetitive coding-testing-debugging development cycles. \pyRobots
provides several tools to support this development phase. A special attention
was paid to the logging infrastructure. It relies on the versatile Python
built-in logging tools, which support configurable parallel logging streams
(typically, a full debug stream is logged to a file while a less verbose stream
is output on the console at run-time).

\pyRobots logs are designed to allow full \emph{post-mortem} replay, and we
provide a graphical interface to display the actions and resource usage over
time (Figure~\ref{log_view}).

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{log}
        \caption{Screenshot of \pyRobots log analyser, processing the log of the
        experiment presented at section~\ref{croq}}
        \label{log_view}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Detailed Example}
% 
% \begin{listing}[h!]
% 
% \begin{pythoncode}
%     import time
%     from robots import GenericRobot
%     from robots.decorators import action, lock
%     from robots.resources import Resource
%     from robots.signals import ActionCancelled
% 
%     # create a 'lockable' resource for our robot
%     WHEELS = Resource("wheels")
% 
%     class MyRobot(GenericRobot):
% 
%       def __init__(self):
%         super(MyRobot, self).__init__()
% 
%         # create (and set) one element in the robot's state.
%         # Here a sonar.
%         self.state["sonar"] = 10.
% 
%         # do whatever other initialization you need :-)
% 
%       def send_goal(self, pose):
%         # move your robot using your favorite middleware
%         print("Starting to move towards %s" % pose)
% 
%       def stop(self):
%         # stop your robot using your favorite middleware
%         print("Motion stopped")
% 
%       def whatever_lowlevel_method_you_need(self):
%         pass
% 
%     @lock(WHEELS)
%     @action
%     def move_forward(robot):
%       """ Actions are written in a simple imperative, 
%           blocking style.
%       """
% 
%       # the target pose: simply x += 1.0m in the robot's 
%       # frame. pyRobots will handle the frames 
%       # transformations as needed.
%       target = [1.0, 0., 0., "base_link"]
% 
%       try:
%         robot.send_goal(target)
% 
%         while(robot.pose.distance(robot.pose.myself(), 
%                                   target) > 0.1):
%             # robot.sleep is exactly like time.sleep, 
%             # except it lets the pyrobots signals pass 
%             # through.
%             robot.sleep(0.5)
% 
%         print("Motion succeeded")
% 
%       except ActionCancelled:
%         # if the action is cancelled, clean up your state
%         robot.stop()
% 
% 
%     with MyRobot() as robot:
% 
%       # Turn on DEBUG logging.
%       # Shortcut to configure the std Python logging system
%       robot.debug()
% 
%       robot.whenever("sonar", below = 0.3).do(move_forward)
% 
%       try:
%         while True:
%           time.sleep(0.5)
%       except KeyboardInterrupt:
%         pass
% \end{pythoncode}
% \caption{Full example of a \pyRobots application.}
% \label{mwe}
% \end{listing}
%

@@

The development of \pyRobots started in 2011, driven by the requirements of a
18-min long theatre performance between a human actor and a robot. Since that time, the toolkit has
iteratively matured, and has been deployed and used on several projects and
platforms. We first briefly recap these deployments with references to the
corresponding publications, and then present our latest study, in a nursery,
that recently acted as a stress-test for the library.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
\begin{center}
\begin{tabular}{p{0.9\columnwidth}}
    \toprule
    {\bf Manipulation} \\
     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
     \midrule
    {\bf Gaze control} \\
     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
     \midrule
    {\bf Navigation} \\
     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
     \midrule
    {\bf Local navigation} \\
     {\tt dock}, {\tt rotate}, {\tt translate} \\
     \midrule
    {\bf Posture configuration} \\
     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
     \bottomrule
\end{tabular}
\end{center}
\caption{Examples of \pyRobots actions, mostly extracted from experiments involving
    the PR2 robot.}

@@

\paragraph{Roboscopie} The first complex deployment scenario for \pyRobots was
indeed a theatre performance entitled
\emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The toolkit was created out of
the need to control a complex robot (a Willow Garage PR2) in a structured way,
as required for a live performance with a human\footnote{The original \pyRobots
script used for the performance can be accessed
from~\url{http://www.laas.fr/roboscopie}. Note that this script does not
reflect anymore the toolkit usage.}

A total of 47 \pyRobots actions were written for this event
(Table~\ref{pyrobots_actions} gives a few examples), and the toolkit proved to
be especially convenient to interleave calls to both ROS-based and {\sc
GenoM}-based\footnote{{\sc GenoM} is software component as well as a middleware
developed at LAAS-CNRS} components through their respective Python bindings.

\pyRobots also allowed us to retain the \emph{rapid prototyping} convenience of a
script language, essential for the robot to ``practise'' and adapt to the
vision of the director.

@@

\paragraph{Stress-test in a nursery}
\label{croq}

@@

Software-wise, all the user-facing behaviours are implemented with \pyRobots,
and the communication with the hardware is abstracted through a mix of ROS
and Aseba~\cite{magnenat2011aseba} nodes.

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviour of the robot is controlled with \pyRobots.}
        \label{expe-nursery}
\end{figure}

@@

        robot.every("pacifier", becomes = True).do(on_pacifier)
        robot.every("pacifier", becomes = False)
                                       .do(on_pacifier_removed)
        robot.every("scale", increase = 0.3).do(on_toy)
        robot.every("bumper", becomes = True).do(on_bumper)

>>>>

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\pyRobots}{\textsc{pyRobots}}

\title{\LARGE \bf
    \pyRobots{}, a toolset for robot executive control
}

@@

Presented is \pyRobots{}, a new software toolkit for the executive control of
complex robotic systems. Borrowing ideas from previous tools like {\sc
urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
develop (pre-emptive) concurrent and event-based executive controllers.

Designed in a \emph{bottom-up} fashion, out of the need for a practical,
unobstrusive tool suitable for rapid prototyping of complex interaction
scenarios, \pyRobots{} also exposes several simple yet convenient abstractions for
physical resources management and pose representation.

@@

Orchestrating the activity of a robot is a central issue of robotics: \emph{what
to do when?} Traditionally, in layered architectures, this function is
implemented in the so-called \emph{deliberative layer} where decisions are made
based on perceptions and on the current internal state, and executed by sending
orders to a \emph{functional layer}.

@@

This article introduces our attempt at designing an unobtrusive execution
control toolkit, that aims at addressing the \emph{practicality} issue: it has
been designed and implemented from the bottom-up, starting from actual needs
when running large robotic systems (including Willow Garage's PR2 and
Aldebaran's Nao) in dynamic and partially unpredictable environments (typically,
loosely constrained human-robot interaction scenarios).

Instead of an \emph{environment} or a \emph{framework}, which would suggest
rigid design and development guidelines, \pyRobots{} can be conceived as a set
of software helpers to write parallel, event-based, high-level robot
controllers.  It is written for and in Python, which ensure familiarity, fast
development cycles and broad support for interfacing with existing robots and
middlewares. \pyRobots{} is indeed \emph{not} a middleware: it purposefully
provides no mechanisms to connect to other modules or components. It instead
\emph{relies} on one (or several) middleware to actually communicate with the
robot and perform actions.

Also, \pyRobots{} does not provide formal models or verifiability: in its
current state, it does not attempt to contribute in this field. It focuses
instead on providing a lightweight tool to control robots in environments that
require carrying out many tasks in parallel, in an unpredictable, event-based
manner. As such, \pyRobots{} is focusing on use-cases ranging from prototyping
to mid-size experiments.

@@

Any robotic experiment requires a certain amount of high-level control to
supervise the robot behaviour, and, expectedly, many projects have attempted to
design and provide tools for this task. Degroote provides an in-depth review of
the existing techniques and tools in~\cite{degroote2012architecture}. We focus
this review on tools dedicated to the explicit implementation of high-level
behaviours, excluding complete control stacks (that would typically include a
middleware and hardware abstractions), meta-tools (like component generators or
model checkers) and (semi-)automatic techniques (like emerging behaviours or
learning by demonstration).  In contrast with middlewares for which the robotic
community relies on a few \emph{de-facto} standards (ROS, YARP, OpenRTM), we
want to underline that no such widely-used softwares exist for execution
control, and we believe that most of today's robot behaviours are implemented as
\emph{ad-hoc} scripts.

@@

Domain-specific languages (DSL), in the widest sense, represent the largest body
of literature on high-level behaviour programming. They can either be ``actual''
domain-specific languages like PRS~\cite{ingrand1996prs},
RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
derivatives like {\sc ReadyLOG}~\cite{ferrein2007robot}) or, more recently,
{\sc urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
languages like Python or Ruby, or logic languages like Prolog) to provide
high-level paradigms and control structures well suited to the programming of
robots. {\sc cram}~\cite{beetz2010cram} (based on Lisp and Prolog), {\sc
roby}~\cite{joyeux2009plan} (based on Ruby), {\sc
teer}~\cite{magnenat2012teer} (based on Python), are
examples of such eDSL.

@@

\pyRobots{} positions itself in this landscape as a lightweight eDSL (in a similar
fashion to {\sc roby} or {\sc teer}), that focuses on high-level behaviours
implementation. It seeks unobstrusive interaction with the other components of
the robot software stack, using Python as ubiquitous binding language. Similar
to {\sc urbi}, it focuses on transparent preemptive concurrency and event-based programming.
Also, \pyRobots{}' model favours hierarchical task composition over plans, and does not
provide by itself any task planning service, delegating this to external tools.

@@

This section introduces the main concepts of \pyRobots{},
and provide an overview of the features and mechanisms. Since \pyRobots{} is an
open-source software, the interested reader can also refer to the source code
and documentation, available from~\url{https://github.com/chili-epfl/pyrobots}.

@@

\pyRobots{} core concepts are simple: on one hand, the user creates a
\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
different low-level \textbf{controllers} (proxies to actuators, typically
provided by a middleware like ROS) and the \textbf{state} of the robot
(typically a set proxies to the sensors, or a connector to a database, etc.,
accessed through a key-value container -- a dictionary-like object).

On the other hand, the user creates as many high-level \textbf{actions} as
desired, as regular Python functions, only annotated with the \python|@action|
decorator. Actions are automatically added to the
\textbf{robot}, and can access its state and controllers to perform actual
physical actions.

@@

Finally, the user can declare arbitrary \textbf{resources}, usually
corresponding to the robot physical components (for instance, the wheels, the
camera) and guarantee exclusive access to the required resource to specific
actions.

@@

\emph{Actions} in \pyRobots{} are asynchronous by default. They are implemented as
\emph{futures} (also known as \emph{promises}) and therefore executed in
independent threads. To the developer, however, they appear as regular Python
function annotated with the \python|@action| decorator, and are invoked
like any other function.  Since they are asynchronous, these functions may
perfectly block or even never return (typically useful to implement continuous
background behaviours).

@@

\emph{Signaling:} Behaviours often need to be paused or cancelled.  This raises
an issue that specifically needs to be addressed in robotics: before suspending
an action, one often wants to first bring back the robot into some form of rest
state (\eg one would not want to suddenly cancel a \python|walk| action in the
middle of a step without first bringing back the foot to the ground). \pyRobots{}
addresses this issue through signals, and extends standard Python thread into
\emph{signaling threads}, relying on Python's usual syntax for exception
handling. Listing~\ref{lst:signals} illustrates this mechanism. The action is
instantiated and started at line 8, and cancelled at line 10. This send a
\python|ActionCancelled| signal to the action that the user can handle using an
\python|except| statement.

\begin{listing}[H]
\begin{pythoncode}
    @action
    def safe_walk(robot):
      try:
        robot.walk()
      except ActionCancelled:
        robot.go_back_to_rest_pose()

    action = robot.safe_walk()
    time.sleep(1)
    action.cancel()
\end{pythoncode}
\caption{\textbf{Handling a cancellation signal} After one second, the
\python|safe_walk| action is cancelled. This sends the signal
\python|ActionCancelled| to the action, that can appropriately terminate.}
\label{lst:signals}
\end{listing}

Pausing actions is more complex (the action internal state must be saved, a mean
to resume the action where it was interrupted must be provided, the user must be
given ways to code transition behaviours when pausing and resuming) and is not
currently supported by \pyRobots{}.

@@

Sharing the same syntax with {\sc urbi}, events in \pyRobots{} can by monitored
through the methods \python|on| (one shot event) and \python|whenever|
(continuous event monitoring). Event conditions can be either a predicate (that
must take as single argument the robot instance) or a pair of $\{\text{state
key}, \text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$.
Examples of event usage can be seen at lines 33--38 of
listing~\ref{croquignole_no_move}.

Each event monitor lives in its own separate thread, and when the event is fired,
the callback (usually, an action) is invoked from the event monitor thread
itself.

@@

Physical robot resources (mainly actuators, but possibly sensors as
well) usually mandate exclusive use (it is seldom desirable that two different actions
control the wheels of a robot at the same time). \pyRobots{} provides a resource
locking mechanism that relies on the \python|@lock| annotation: an action
annotated with \python|@lock(ARMS)| (Listing~\ref{lst:resources}, l.6) will
acquire (or wait for) the \python|ARM| resource when being executed.

This mechanism is generic and \pyRobots{} makes no assumptions concerning the
actual resources available/meaningful to a specific robot. The
programmer define them himself, either as stand-alone or as compound
resources (listing~\ref{lst:resources}, l.1--3).

@@

    robot.lift_box()
    robot.wave_hand() # waits until lift_box is over
    robot.scratch_head() # skipped if lift_box or
                         # wave_hand are still running

\end{pythoncode}
\caption{\textbf{Resource locking} Resource usage is defined at the
action-level, through annotations.}

\label{lst:resources}
\end{listing}

A given action can lock several distinct resources (with the corresponding
number of \python|@lock| annotations. When executed, the
action acquires a lock on the resource (and all dependent resources in the case
of a compound resource) which is released upon action completion. If the lock is
already acquired, two strategies are possible: either wait or \emph{fail fast}
(\python|@lock(..., wait=False)|, l.16).

@@

\pyRobots{} offers a similar approach \emph{within} the control script, adding a
pinch of syntactic sugar to make poses and frame management transparent for the
high-level action controller.

As illustrated in listing~\ref{lst:poses}, poses can be flexibly defined by a
simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt map} is then
assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y, z, rx, ry,
rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx, ry, rz,
frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x, y, z,
qx, qy, qz, qw, frame\}$ or by a dictionary that defines any of these 8 values.
Internally, \pyRobots{} normalizes and stores the poses in their full 8-tuple
expansion.

@@

\end{pythoncode*}
\caption{Examples of \textbf{pose normalization}. Poses can be transformed to
other reference frames with the \python|inframe| method (which implicitly performs
normalization if needed).}
\label{lst:poses}
\end{listing}

@@

These transformations are directly computed, but \pyRobots{} also takes advantage
of middleware-specific infrastructure (when available) to provide known frames.
ROS TF, Aldebaran's {\tt NAOqi} frame conventions, etc. are used when available
through specializations of the \pyRobots{} \python|PoseManager| class.

@@

Running robots in real-world, \ie dynamic and unpredictable, environments
entails repetitive coding-testing-debugging development cycles. \pyRobots{}
provides several tools to support this development phase. A special attention
was paid to the logging infrastructure. It relies on the versatile Python
built-in logging tools, which support configurable parallel logging streams
(typically, a full debug stream is logged to a file while a less verbose stream
is output on the console at run-time).

\pyRobots{} logs are designed to allow full \emph{post-mortem} replay, and we
provide a graphical interface to display the actions and resource usage over
time (Figure~\ref{log_view}).

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{log}
        \caption{Screenshot of \pyRobots{} log analyser, processing the log of the
        experiment presented at section~\ref{croq}}
        \label{log_view}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Detailed Example}
% 
% \begin{listing}[h!]
% 
% \begin{pythoncode}
%     import time
%     from robots import GenericRobot
%     from robots.decorators import action, lock
%     from robots.resources import Resource
%     from robots.signals import ActionCancelled
% 
%     # create a 'lockable' resource for our robot
%     WHEELS = Resource("wheels")
% 
%     class MyRobot(GenericRobot):
% 
%       def __init__(self):
%         super(MyRobot, self).__init__()
% 
%         # create (and set) one element in the robot's state.
%         # Here a sonar.
%         self.state["sonar"] = 10.
% 
%         # do whatever other initialization you need :-)
% 
%       def send_goal(self, pose):
%         # move your robot using your favorite middleware
%         print("Starting to move towards %s" % pose)
% 
%       def stop(self):
%         # stop your robot using your favorite middleware
%         print("Motion stopped")
% 
%       def whatever_lowlevel_method_you_need(self):
%         pass
% 
%     @lock(WHEELS)
%     @action
%     def move_forward(robot):
%       """ Actions are written in a simple imperative, 
%           blocking style.
%       """
% 
%       # the target pose: simply x += 1.0m in the robot's 
%       # frame. pyRobots will handle the frames 
%       # transformations as needed.
%       target = [1.0, 0., 0., "base_link"]
% 
%       try:
%         robot.send_goal(target)
% 
%         while(robot.pose.distance(robot.pose.myself(), 
%                                   target) > 0.1):
%             # robot.sleep is exactly like time.sleep, 
%             # except it lets the pyrobots signals pass 
%             # through.
%             robot.sleep(0.5)
% 
%         print("Motion succeeded")
% 
%       except ActionCancelled:
%         # if the action is cancelled, clean up your state
%         robot.stop()
% 
% 
%     with MyRobot() as robot:
% 
%       # Turn on DEBUG logging.
%       # Shortcut to configure the std Python logging system
%       robot.debug()
% 
%       robot.whenever("sonar", below = 0.3).do(move_forward)
% 
%       try:
%         while True:
%           time.sleep(0.5)
%       except KeyboardInterrupt:
%         pass
% \end{pythoncode}
% \caption{Full example of a \pyRobots{} application.}
% \label{mwe}
% \end{listing}
%

@@

The development of \pyRobots{} started in 2011, driven by the requirements of a
18-min long theatre performance between a human actor and a robot. Since that time, the toolkit has
iteratively matured, and has been deployed and used on several projects and
platforms. We first briefly recap these deployments with references to the
corresponding publications, and then present our latest study, in a nursery,
that recently acted as a stress-test for the library.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
\begin{center}
\begin{tabular}{p{0.9\columnwidth}}
    \toprule
    {\bf Manipulation} \\
     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
     \midrule
    {\bf Gaze control} \\
     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
     \midrule
    {\bf Navigation} \\
     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
     \midrule
    {\bf Local navigation} \\
     {\tt dock}, {\tt rotate}, {\tt translate} \\
     \midrule
    {\bf Posture configuration} \\
     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
     \bottomrule
\end{tabular}
\end{center}
\caption{Examples of \pyRobots{} actions, mostly extracted from experiments involving
    the PR2 robot.}

@@

\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
indeed a theatre performance entitled
\emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The toolkit was created out of
the need to control a complex robot (a Willow Garage PR2) in a structured way,
as required for a live performance with a human\footnote{The original \pyRobots{}
script used for the performance can be accessed
from~\url{http://www.laas.fr/roboscopie}. Note that this script does not
reflect anymore the toolkit usage.}

A total of 47 \pyRobots{} actions were written for this event
(Table~\ref{pyrobots_actions} gives a few examples), and the toolkit proved to
be especially convenient to interleave calls to both ROS-based and {\sc
GenoM}-based\footnote{{\sc GenoM} is software component as well as a middleware
developed at LAAS-CNRS} components through their respective Python bindings.

\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience of a
script language, essential for the robot to ``practise'' and adapt to the
vision of the director.

@@

\subsubsection{Stress-test in a nursery}
\label{croq}

@@

Software-wise, all the user-facing behaviours are implemented with \pyRobots{},
and the communication with the hardware is abstracted through a mix of ROS
and Aseba~\cite{magnenat2011aseba} nodes.

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviour of the robot is controlled with \pyRobots{}.}
        \label{expe-nursery}
\end{figure}

@@

        robot.whenever("pacifier", becomes = True)
                                    .do(on_pacifier)
        robot.whenever("pacifier", becomes = False)
                                    .do(on_pacifier_removed)
        robot.whenever("scale", increase = 0.3).do(on_toy)
        robot.whenever("bumper", becomes = True).do(on_bumper)
